Efficient pedigree recording for fast population genetics simulation

### 2017-11-16 forwards_paper.tex
moved API to methods

\plr{leaving this alone for the moment: to stay here? move?}

The \msprime\ Python API provides a powerful platform for working with
tree topology and mutation data. We refer to the part of \msprime\
that is dedicated to tree sequence input and output as the `Tables API',
as the API is organised around simple tables of data. There are four key
tables: nodes, edges, sites and mutations. Briefly, nodes and edges define
the topology of a tree sequence (as defined above) and the sites and mutations
define mutational processes on this topology.
Figure~\ref{fig:example_tree_sequence} gives an informal depiction of this
encoding in terms of these tables.

The tables API is primarily designed to facilitate efficient interchange of
data between programs or between different modules of the same program.
Following the current best-practises [citations: apache arrow, etc] data is stored
in a columnar format. There are many advantages to this approach, but the
principle advantage for our purposes is that it allows for very efficient
interchange of large amounts of numerical data. In principle, this enables
zero-copy semantics, where a data consumer can read the information directly
from the memory of a producer without incurring the overhead of a copy
[citation?] Our implementation uses the numpy C API [citation] to efficiently copy
data from Python into the low-level C library used to manipulate
tree sequences.

The tables API provides basic input and output operations via the numpy
array interface, which provides a great deal of flexibility as well
as efficiently. This makes transferring data from various sources
such as HDF5 [citation], Dask, Zarr etc, straightforward. (For small
scale data and debugging purposes, a simple text format is also supported.)
Along with
these operations we also provide a function to sort a set of tables
to ensure that the records are in the form required to input
into an \msprime\ tree sequence object. The \texttt{simplify\_tables}
function implements tree sequence simplification, as described in the
previous subsection.

This interchange API is very efficient. [Describe a quick example where we generate
a many-gigabyte tree sequence using fwdpp, and the time required
to copy the node and edge data into the tables]. By using a simple numerical
encoding of tree topologies and contiguous arrays of data to store this
encoding, we can achieve data transfer rates that would be impossible under
any text-based approach while retaining excellent portability.

\subsection*{Estimates of run-time complexity}

@@

%%%%%%%%%%%%%%
\subsection*{Data structures:}
\jk{Moved from above.}

>>>>

%%%%%%%%%%%%
\subsection*{Estimates of run-time complexity}

@@

The \msprime{} Python API provides a powerful platform for working with tree topology and mutation data. 
The new portions of \msprime{} that we cover here
are dedicated to tree sequence input and output using simple tables of data,
so we refer to this as the `Tables API'.
The four key tables: nodes, edges, sites and mutations, are described above.

The tables API is primarily designed to facilitate efficient interchange of
data between programs or between different modules of the same program.
Following the current best-practises [citations: apache arrow, etc] data is stored
in a columnar format.
The principal advantage of this for our purposes is that it allows for very efficient
interchange of large amounts of numerical data. In principle, this enables
zero-copy semantics, where a data consumer can read the information directly
from the memory of a producer without incurring the overhead of a copy
[citation?] Our implementation uses the numpy C API [citation] to efficiently copy
data from Python into the low-level C library used to manipulate
tree sequences.
\plr{@molpopgen: note on how this works with fwdpp?}

The tables API provides basic input and output operations via the numpy
array interface, which provides a great deal of flexibility as well
as efficiency, and makes it straightforward to transfer data from sources
such as HDF5 [citation], Dask, or Zarr. 
(For small scale data and debugging purposes, a simple text format is also supported.)
Along with these operations we also provide a function to sort a set of tables
to ensure that the records are in the form required to input
into an \msprime{} tree sequence object. (Simplification may also be required.)

\plr{Maybe we don't need this paragraph? Already said it, more or less, above, esp if we insert something about how fwdpp works.}
This interchange API is very efficient. [Describe a quick example where we generate
a many-gigabyte tree sequence using fwdpp, and the time required
to copy the node and edge data into the tables]. By using a simple numerical
encoding of tree topologies and contiguous arrays of data to store this
encoding, we can achieve data transfer rates that would be impossible under
any text-based approach while retaining excellent portability.

%%%%%%%%%%%%%%
\subsection*{Data structures:}
